Context-dependent consequences of including lagged effects in demographic models

### 2023-07-20 docs/rmd_version/paper.Rmd
adding quarto version of ms

# header-includes:
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyfoot{}
  - \fancyhead[R]{p. \thepage}
  - \fancyhead[L]{Draft - `r format(Sys.time(), '%d %B %Y')`}
  - \usepackage{sectsty} \sectionfont{\centering}
  - \usepackage[utf8]{inputenc} # This loads package to convert unicode 
  - \DeclareUnicodeCharacter{394}{$\Delta$} # convert unicode \u0394 to Delta
  # - \usepackage{sectsty} \subsectionfont{\centering}
  # - \AtBeginDocument{\let\maketitle\relax}
 #  - \usepackage{float} #to keep figure in the location placed instead of end
 #  - \floatplacement{figure}{H}
 #  - \usepackage{booktabs}
 #  - \usepackage{setspace} # line spacing
 #  - \raggedbottom
 # # - \usepackage{endfloat} #[nomarkers] excludes the {insert figure x around here] from main text. The others exclude the list of tables and figures. https://cs.brown.edu/about/system/managed/latex/doc/endfloat.pdf
 #  - \usepackage{tabu}
 #  - \usepackage{makecell}
 #  - \usepackage{pdflscape}
 #  - \usepackage{longtable}
 #  - \newcommand{\blandscape}{\begin{landscape}}
 #  - \newcommand{\elandscape}{\end{landscape}}
 #  - \setlength\parindent{22pt}
---

@@

@@

2.  It has long been recognized that there is the potential for lagged effects.
    Including lagged effects in models has been a major technical challenge, but there are now multiple approaches for doing so.
    The studies assessing the potential for lagged effects on vital rates find that they indeed appear to be prevalent [@eversLaggedDormantSeason2021; @scottDelayedEffectsClimate2022].

3.  Lagged effects could be different in different habitat types (and this could be a big reason for differences between habitats)

# Methods

## Demographic Data

<!--# Emilio will fill this section -->

## IPMs

@@

# Results

For all vital rates estimated using the long term demographic dataset, the DLNM model fit the best (dAIC = 0) followed by the model with a random effect of year, followed by the deterministic model (Table \@ref(tab:aic)).
<!--# No longer true if random effect = different smooth for each year -->

@@

>>>>

# header-includes:
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyfoot{}
  - \fancyhead[R]{p. \thepage}
  - \fancyhead[L]{Draft - `r format(Sys.time(), '%d %B %Y')`}
  - \usepackage{sectsty} \sectionfont{\centering}
  - \usepackage[utf8]{inputenc} # This loads package to convert unicode 
  - \DeclareUnicodeCharacter{394}{$\Delta$} # convert unicode \u0394 to Delta
  # more info # https://www.ctan.org/pkg/sectsty
  - \usepackage{sectsty} \sectionfont{\centering}
  # to center to sub-sub-headings (###) and change color of font
  - \usepackage{sectsty} \subsectionfont{\centering}
  # - \usepackage{sectsty} \subsubsectionfont{\centering}
  # - \usepackage{sectsty} \subsectionfont{\centering}
  # - \AtBeginDocument{\let\maketitle\relax}
 #  - \usepackage{float} #to keep figure in the location placed instead of end
 #  - \floatplacement{figure}{H}
 #  - \usepackage{booktabs}
 #  - \usepackage{setspace} # line spacing
 #  - \raggedbottom
 # # - \usepackage{endfloat} #[nomarkers] excludes the {insert figure x around here] from main text. The others exclude the list of tables and figures. https://cs.brown.edu/about/system/managed/latex/doc/endfloat.pdf
 #  - \usepackage{tabu}
 #  - \usepackage{makecell}
 #  - \usepackage{pdflscape}
 #  - \usepackage{longtable}
 #  - \newcommand{\blandscape}{\begin{landscape}}
 #  - \newcommand{\elandscape}{\end{landscape}}
 #  - \setlength\parindent{22pt}
---

@@

1.  **Paragraph 1**

@@

2.  **Paragraph 2**

    1.  It has long been recognized that there is the potential for lagged effects.\
    2.  Including lagged effects in models has been a major technical challenge, but there are now multiple approaches for doing so.\
    3.  The studies assessing the potential for lagged effects on vital rates find that they indeed appear to be prevalent [@eversLaggedDormantSeason2021; @scottDelayedEffectsClimate2022].

3.  **Paragraph 3**

    1.  Lagged effects could be different in different habitat types.\
    2.  This could be a big reason for differences between habitats.

4.  **Paragraph 4**

    1.  Here we...
    2.  We conducted these analyses with both deterministic and stochastic IPMs.

## Methods

### Demographic methods and data

Overviw of <!--# Emilio will fill this section -->

<!-- Add info about additional datasets (Emilio) -->

### Construction of Integral Projection Models

@@

### Statistical analyses

All about the stats.

## Results

For all vital rates estimated using the long term demographic dataset, the DLNM model fit the best (dAIC = 0) followed by the model with a random effect of year, followed by the deterministic model (Table \@ref(tab:aic)).\
<!--# No longer true if random effect = different smooth for each year -->

@@

```{r time, results='asis'}
meta_df <- 
  tar_meta(store = here("_targets")) %>% 
  select(name, size, bytes, time, seconds) %>% 
  filter(str_detect(name, "^ipm_\\w+_[cf]{2}$")) %>% 
  mutate(minutes = seconds / 60,
         hours = minutes / 60)

meta_df_tbl<-meta_df %>% 
  separate(name, into = c("trash", "IPM", "Habitat")) %>% 
  select(-trash) %>% 
  select(IPM, Habitat, minutes) %>% 
  group_by(IPM) %>% 
mutate(IPM = str_replace_all(
    IPM,
    c(
      "det" = "Deterministic",
      "stoch" = "Stochastic, kernel-resampled",
      "dlnm" = "Stochastic, parameter-resampled"
    ))) %>%
  summarize(mean_time_min = round(mean(minutes), 2)) %>%
  rename("IPM Type" = IPM) %>% 
  rename("mean time (min.)" = mean_time_min)

kbl(meta_df_tbl, 
    booktabs = T,
    align = "lc",
    caption = "table caption") %>% 
  kable_styling(full_width = F,
                position = "center")
